---
title: "Activity-Dosing-Plot"
author: "Pooja Pednekar"
date: "2026-04-03"
output: html_document
---
This script visualizes all recorded activities and the alkalinity dosing volumes added to the blue tanks and header tank in the CBLS Wet Lab for the Putnam Lab. It should be run daily after measurements have been collected and entered into the datasheet.

Each day, be sure to update the datasheet with any activities performed and any changes to dosing volumes in the “Activity” and “Dosing_mL” columns. The dosing volume in this column represents the volume administered at a single dosing interval. To calculate the total daily dosing volume, multiply this value by 4

The output generated by this script helps us interpret daily measurement trends and assess whether observed changes in water chemistry may be associated with specific activities or dosing adjustments.

Script last updated and uploaded to CBLS Wetlab > Scripts 02/13/2025 by PP.


```{r}
library(tidyverse)
library(lubridate) # used for converting 8 digit date into datetime format for R
library(ggrepel)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
daily <- read.csv("../Data/Daily_measurements_tracking.csv")
head(daily)
```


```{r}
tail(daily) # check to make sure data from today is there
```

```{r}
## Swap blue_tank1 and 2 for BT 1/2
daily$Tank_ID <- ifelse(daily$Tank_ID == "blue_tank1" | daily$Tank_ID == "blue_tank2", "BT 1/2", daily$Tank_ID)

## Swap blue_tank3 and 4 for BT 3/4
daily$Tank_ID <- ifelse(daily$Tank_ID == "blue_tank3" | daily$Tank_ID == "blue_tank4", "BT 3/4", daily$Tank_ID)

## Swap quarantine_tank for QT
daily$Tank_ID <- ifelse(daily$Tank_ID == "quarantine_tank", "QT", daily$Tank_ID)
```

#Filtering data based on the columns needed and converting date format
```{r}
activity_tanks <- daily %>%
  select(Date, Tank_ID, Activity, Alk_Dosing_ml, notes) %>%
  mutate(
    Date = ymd(as.character(Date)),             #date format fixture
    Activity = na_if(str_trim(Activity), ""))   #empty colons resembling no activity 
```


#plotting only for past three months; could be changed based on need
```{r}
activity_month <- activity_tanks %>%
  filter(Date >= Sys.Date() - 60) #change the number of days based on your consideration
```

#creating dataframe for activity panel (into binary) for the final plot
```{r}
#Plotting only data when the activity occurred
activity_yes <- activity_month %>%
  filter(!is.na(Activity)) %>%
  distinct(Date, Activity,Tank_ID) %>%
  mutate(
    metric = "Activity",
    value  = ifelse(Tank_ID=="BT 1/2",1,
                    ifelse(Tank_ID=="BT 3/4", 2,
                           3))
  )

#activity_plot_df <- bind_rows(activity_yes) %>%  #useful if we also want to plot the no activity data points in future (not needed for our plot)
 # mutate(
  #  metric = factor(metric, levels = c("Activity", "Dosing")),
   # Activity = fct_infreq(Activity)
  #)
```

#creaing dataframe for dosing panel
```{r}
dosing_plot_df <- activity_month %>%
  filter(!is.na(Alk_Dosing_ml), Alk_Dosing_ml > 0) %>%
  mutate(
    metric = factor("Dosing", levels = c("Activity", "Dosing")),
    value  = Alk_Dosing_ml
  )
```

#Final Plot: Activity & Dosing
```{r}
#activity plot is plotted as binary; meaning if activity occurred n=; no activity n=0.
activity_dosing_plot <- ggplot() +
  geom_text(
    data = activity_yes,
    aes(x = Date, y=value, label = Activity, color=Tank_ID),
    size = 2,
    angle= 45,
    hjust=1
  ) +
    geom_point(
    data = activity_yes,
    aes(x = Date, y=value,color=Tank_ID),
    size = 1
  ) +

#The dosing volume data represents the volume administered at a single dosing interval. To calculate the total daily dosing volume, multiply this value by 4 as we are dosing the Ht with alkalinity 4x/day.
  geom_col(
    data = dosing_plot_df,
    aes(x = Date, y = value, fill = Tank_ID),
    position = "dodge",
    width = 0.8
  ) +
  facet_grid(metric ~ ., scales = "free_y",
             labeller= as_labeller(c( Activity= "Activity Timeline", Dosing ="Dosed 4x day (mL)")))+
  xlab("Date") + ylab("value") +
  theme_bw() +
  theme(text = element_text(size = 14),
        strip.text.y = element_text(size = 12)) +
  scale_x_date(date_breaks = "1 day", date_labels = "%a\n%b %d") 


activity_dosing_plot
```


```{r}
#saving output in the output folder
ggsave("../Output/Activity_Dosing_Plot.pdf",
       activity_dosing_plot,width = 20, height = 8, units = "in")
```

