---
title: "Daily_measurements"
author: "LaurenZane"
date: "7/19/2022"
output:
  pdf_document: default
  html_document: default
---

This script plots daily measurements from the CBLS Wet lab for the Putnam lab. Run this script daily after daily measurements are taken and inputted into the datasheet. 

Script updated 11/14/23 by ZD. 

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate) # used for converting 8 digit date into datetime format for R
library(RColorBrewer)
library(rmarkdown)
library(tinytex)

## If seacarb needs to be downloaded:
#packageurl <- "https://cran.r-project.org/src/contrib/Archive/seacarb/seacarb_3.2.tar.gz"
#install.packages(packageurl, repos = NULL, type = "source")
#install.packages("seacarb")
library(seacarb) 
```

## Load data 
```{r}
getwd()
daily <- read.csv("../Data/Daily_measurements_tracking.csv")
head(daily)
tail(daily) # check to make sure data from today is there
```

Set dates as characters (needs to be done for merging with the tris calibration file)
```{r}
daily$Date <- as.character(daily$Date)
daily$tris.date <- as.character(daily$tris.date)
```


Remove quarantine_tank from the dataframe (this line can always be commented out if interested in QT values)

```{r}
daily <- daily %>%
  filter(!Tank_ID == "quarantine_tank") %>%
  filter(!Tank_ID == "blue_tank1") %>% 
  filter(!Tank_ID == "blue_tank2")

```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV. 
```{r}
pHcalib<-read_csv("../Data/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope <- pHcalib %>%
  group_by(tris.date) %>%
  nest() %>%
  mutate(fitpH = map(data, ~ lm(mVTris ~ TTris, data = .x))) %>%
  mutate(tidy_fit = map(fitpH, broom::tidy)) %>%
  unnest(tidy_fit) %>%
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  left_join(daily, ., by = "tris.date") %>%
  mutate(mVTris = Temperature_C * TTris + `(Intercept)`) %>%
  filter(!is.na(mVTris) & !is.na(Temperature_C) & !is.na(Salinity_psu) & !is.na(pH_mv)) %>%
  mutate(pH.total = seacarb::pH(Ex = pH_mv, Etris = mVTris, S=Salinity_psu, T=Temperature_C))
```

The code above appears to not want to run when there are so many NAs in the data, but that becomes a problem when there is so many NAs in the temp and salinity columns. 

Convert date to ymd for plotting 
```{r}
pHSlope$Date <- ymd(pHSlope$Date) # convert 8 digit date into datetime format

pHSlope <- pHSlope %>% relocate("pH.total", .after = Salinity_psu)
```

## Change to long format

Change data format to long format 
```{r}
pHSlope.long <-pHSlope %>% pivot_longer(cols=Temperature_C:pH.total,
  names_to = "metric",
  values_to = "value")
```

Filter by relevant dates if needed

## Plot

First, plot all dates 
```{r}
daily_tank<-pHSlope.long %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw() +
  theme(text = element_text(size = 14)); daily_tank

# Save plot 
ggsave("../Output/Daily_Measurements.pdf", daily_tank, width = 20, height = 15, units = c("in"))
```

It isn't super informative to look at multiple years in one plot. 

Filter and plot by the past month
```{r}
# Filter to the past 30 days 
daily_month <- pHSlope.long %>%
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

# Plot data from the past month
daily_month<-daily_month %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw() +
  theme(text = element_text(size = 14)); daily_month

# Save plot 
ggsave("../Output/Daily_Measurements_Past_Month.pdf", daily_month, width = 20, height = 15, units = c("in"))
```

Summarize daily measurements over the past month (this will use the `daily` df from above instead of the the `daily.long` df)
```{r}
daily_month <- pHSlope %>% 
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

summary<-daily_month%>%
  group_by(Tank_ID)%>%
  #select(!tank)%>%
  select(!Date)%>%
  select(!light2:notes)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary
```

## BT 1 and 2 values

```{r}
getwd()
daily <- read.csv("../Data/Daily_measurements_tracking.csv")
head(daily)
tail(daily) # check to make sure data from today is there
```

Set dates as characters (needs to be done for merging with the tris calibration file)
```{r}
daily$Date <- as.character(daily$Date)
daily$tris.date <- as.character(daily$tris.date)
```


Remove quarantine_tank from the dataframe (this line can always be commented out if interested in QT values)

```{r}
daily_BT1_2 <- daily %>%
  filter(!Tank_ID == "quarantine_tank") %>%
  filter(!Tank_ID == "header_tank") %>%
  filter(!Tank_ID == "blue_tank3") %>% 
  filter(!Tank_ID == "blue_tank4")

```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV. 
```{r}
pHcalib_BT1_2<-read_csv("../Data/Tris_Calibration_BT1_2.csv")
pHcalib_BT1_2$tris.date<-as.character(pHcalib_BT1_2$tris.date)

pHSlope_BT1_2 <- pHcalib_BT1_2 %>%
  group_by(tris.date) %>%
  nest() %>%
  mutate(fitpH = map(data, ~ lm(mVTris ~ TTris, data = .x))) %>%
  mutate(tidy_fit = map(fitpH, broom::tidy)) %>%
  unnest(tidy_fit) %>%
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  left_join(daily, ., by = "tris.date") %>%
  mutate(mVTris = Temperature_C * TTris + `(Intercept)`) %>%
  filter(!is.na(mVTris) & !is.na(Temperature_C) & !is.na(Salinity_psu) & !is.na(pH_mv)) %>%
  mutate(pH.total = seacarb::pH(Ex = pH_mv, Etris = mVTris, S=Salinity_psu, T=Temperature_C))
```

```{r}
pHSlope_BT1_2$Date <- ymd(pHSlope_BT1_2$Date) # convert 8 digit date into datetime format

pHSlope_BT1_2 <- pHSlope_BT1_2 %>% relocate("pH.total", .after = Salinity_psu)
```

## Change to long format

Change data format to long format 
```{r}
pHSlope_BT1_2.long <- pHSlope_BT1_2 %>% pivot_longer(cols=Temperature_C:pH.total,
  names_to = "metric",
  values_to = "value")
```


```{r}
pHSlope.long <- rbind(pHSlope.long, pHSlope_BT1_2.long)

pHSlope <- rbind(pHSlope, pHSlope_BT1_2)
```


First, plot all dates 
```{r}
daily_tank<-pHSlope.long %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw() +
  theme(text = element_text(size = 14)); daily_tank

# Save plot 
ggsave("../Output/Daily_Measurements.pdf", daily_tank, width = 20, height = 15, units = c("in"))
```


```{r}
# Filter to the past 30 days 
daily_month <- pHSlope.long %>%
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

# Plot data from the past month
daily_month<-daily_month %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw() +
  theme(text = element_text(size = 14)); daily_month

# Save plot 
ggsave("../Output/Daily_Measurements_Past_Month.pdf", daily_month, width = 20, height = 15, units = c("in"))
```


```{r}
daily_month <- pHSlope %>% 
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

summary<-daily_month%>%
  group_by(Tank_ID)%>%
  #select(!tank)%>%
  select(!Date)%>%
  select(!light2:notes)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary
```