---
title: "Daily_measurements"
author: "PutnamLab"
date: "7/19/2022"
output:
  html_document: default
  pdf_document: default
---

This script plots daily measurements from the CBLS Wet lab for the Putnam lab. Run this script daily after daily measurements are taken and inputted into the datasheet. 

Script updated 11/14/23 by ZD. 
Script updated 11/16/23 by JA - added upper and lower threshold lines on plots. 

Script updated 1/2/24 by JA. Added probe set to daily df so that pH is calculated based on the probe set. The data is then merged back together for the plots.

Script updated 3/11/24 by DMB. Added total alkalinity plots and upper and lower threshold lines for TA.

Script updated 9/14/24 by ZD to set Probe Set #2 as probe set for QT.

Script updated 1/30/25 by ZD to add Calcium DMs.
Script updated 3/6/25 by ZD to add Alkalinity dkH measurements.

Script updated 4/8/26 by ZD to include nutrient measurements.

Script updated 5/29/26 by PP to include PAM Measurements. 

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate) # used for converting 8 digit date into datetime format for R
library(RColorBrewer)
library(rmarkdown)
library(tinytex)
library(ggrepel)
library(janitor)
library(lmerTest)
library(ggpubr)
library(lme4)
library(emmeans)


## If seacarb needs to be downloaded:
#packageurl <- "https://cran.r-project.org/src/contrib/Archive/seacarb/seacarb_3.2.tar.gz"
#install.packages(packageurl, repos = NULL, type = "source")
#install.packages("seacarb")
library(seacarb) 
```

## Load data 
```{r}
## Read in data
daily <- read.csv("../Data/Daily_measurements_tracking.csv")
head(daily)
tail(daily) # check to make sure data from today is there

## Swap blue_tank1 and 2 for BT 1/2
daily$Tank_ID <- ifelse(daily$Tank_ID == "blue_tank1" | daily$Tank_ID == "blue_tank2", "BT 1/2", daily$Tank_ID)

## Swap blue_tank3 and 4 for BT 3/4
daily$Tank_ID <- ifelse(daily$Tank_ID == "blue_tank3" | daily$Tank_ID == "blue_tank4", "BT 3/4", daily$Tank_ID)

## Swap quarantine_tank for QT
daily$Tank_ID <- ifelse(daily$Tank_ID == "quarantine_tank", "QT", daily$Tank_ID)

# Load TA measurements data
TA <- read.csv("../Data/CBLS_Aquarium_TA.csv")

# Rename the "SampleID" column to "Tank_ID"
TA.mod <- TA %>%
  rename(Tank_ID = SampleID)

# Filter rows for specific Tank_ID values ("header_tank", "BT 1/2", "BT 3/4")
selected_tanks <- c("header_tank", "BT 1/2", "BT 3/4", "QT")
filtered_data <- TA.mod %>%
  filter(Tank_ID %in% selected_tanks)

# Select only the desired columns
filtered_data_subset <- filtered_data %>%
  select(Date, Tank_ID, TA)
```

Set dates as characters (needs to be done for merging with the tris calibration file)
```{r}
daily$Date <- as.character(daily$Date)
filtered_data_subset$Date <- as.character(filtered_data_subset$Date)
daily$tris.date <- as.character(daily$tris.date)
```

Combine TA data frame and daily measurments
```{r}
daily <- merge(daily, filtered_data_subset, by = c("Date", "Tank_ID"), all.x = TRUE)
```

Probe Set 1 (for the header, all corals, and for the non-quarantine tank)
e.g., BT 3/4 and header tank and salt mixing tank
```{r}
daily.probe1 <- daily %>%
  filter(Probe.Set == "Probe1") 

range(na.omit(daily.probe1$Temperature_C))
range(na.omit(daily.probe1$pH_mv))
range(na.omit(daily.probe1$Salinity_psu))


daily.probe2 <- daily %>%
  filter(Probe.Set == "Probe2") 

range(na.omit(daily.probe2$Temperature_C))
range(na.omit(daily.probe2$pH_mv))
range(na.omit(daily.probe2$Salinity_psu))

```

## Calculate total pH from Probe Set 1 

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV. 
```{r}
pHcalib<-read_csv("../Data/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope.probe1 <- pHcalib %>%
  group_by(tris.date) %>%
  nest() %>%
  mutate(fitpH = map(data, ~ lm(mVTris ~ TTris, data = .x))) %>%
  mutate(tidy_fit = map(fitpH, broom::tidy)) %>%
  unnest(tidy_fit) %>%
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  left_join(daily.probe1, ., by = "tris.date") %>%
  mutate(mVTris = Temperature_C * TTris + `(Intercept)`)

# range(pHSlope.probe1$Temperature_C)
# pHSlope.probe1 <- pHSlope.probe1 %>% filter(!is.na(Temperature_C)) 
# range(pHSlope.probe1$Temperature_C)
# 
# range(pHSlope.probe1$pH_mv)
# pHSlope.probe1 <- pHSlope.probe1 %>% filter(!is.na(pH_mv)) 
# range(pHSlope.probe1$pH_mv)
# 
# range(pHSlope.probe1$Salinity_psu)
# pHSlope.probe1 <- pHSlope.probe1 %>% filter(!is.na(Salinity_psu)) 
# range(pHSlope.probe1$Salinity_psu)
# 
# range(pHSlope.probe1$pH_mv)
# pHSlope.probe1 <- pHSlope.probe1 %>% filter(!is.na(pH_mv)) 
# range(pHSlope.probe1$pH_mv)
# 
# range(pHSlope.probe1$mVTris)
# pHSlope.probe1 <- pHSlope.probe1 %>% filter(!is.na(mVTris)) 
# range(pHSlope.probe1$mVTris)

pHSlope.probe1 <- pHSlope.probe1 %>%
  mutate(pH.total = seacarb::pH(Ex = pH_mv, Etris = mVTris, S=Salinity_psu, T=Temperature_C))
```

## Calculate total pH from Probe Set 2 
Probe Set 2 (for the header, all corals, and for the non-quarantine tank)
e.g., BT 1/2 while HI corals were islotaed and quarantine tank

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV. 
```{r}
pHcalib2<-read_csv("../Data/Tris_Calibration_ProbeSet_2.csv")
pHcalib2$tris.date<-as.character(pHcalib2$tris.date)

pHSlope.probe2 <- pHcalib2 %>%
  group_by(tris.date) %>%
  nest() %>%
  mutate(fitpH = map(data, ~ lm(mVTris ~ TTris, data = .x))) %>%
  mutate(tidy_fit = map(fitpH, broom::tidy)) %>%
  unnest(tidy_fit) %>%
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  left_join(daily.probe2, ., by = "tris.date") %>%
  mutate(mVTris = Temperature_C * TTris + `(Intercept)`)

range(pHSlope.probe2$Temperature_C)
pHSlope.probe2 <- pHSlope.probe2 %>% filter(!is.na(Temperature_C))
range(pHSlope.probe2$Temperature_C)

range(pHSlope.probe2$pH_mv)
pHSlope.probe2 <- pHSlope.probe2 %>% filter(!is.na(pH_mv))
range(pHSlope.probe2$pH_mv)

range(pHSlope.probe2$Salinity_psu)
pHSlope.probe2 <- pHSlope.probe2 %>% filter(!is.na(Salinity_psu))
range(pHSlope.probe2$Salinity_psu)

range(pHSlope.probe2$pH_mv)
pHSlope.probe2 <- pHSlope.probe2 %>% filter(!is.na(pH_mv))
range(pHSlope.probe2$pH_mv)

range(pHSlope.probe2$mVTris)
pHSlope.probe2 <- pHSlope.probe2 %>% filter(!is.na(mVTris))
range(pHSlope.probe2$mVTris)

pHSlope.probe2 <- pHSlope.probe2 %>%
  mutate(pH.total = seacarb::pH(Ex = pH_mv, Etris = mVTris, S=Salinity_psu, T=Temperature_C))

```

Join Probe 1 and Probe 2 Sets
```{r}
pHSlope <- rbind(pHSlope.probe1, pHSlope.probe2)
```

Convert date to ymd for plotting 
```{r}
pHSlope <- pHSlope %>%
  filter(!Tank_ID == "quarantine_tank")

pHSlope$Date <- ymd(pHSlope$Date) # convert 8 digit date into datetime format

pHSlope <- pHSlope %>% relocate("pH.total", .after = Salinity_psu) %>%
  relocate(pH_mv, .after = pH.total)

head(pHSlope)

nutrients <- pHSlope %>% select(Date, Tank_ID, Time, Date.Time, Observer.Initials, Temperature_C,Phosphorus,Nitrite,Ammonia,Probe.Set, notes,P_Dosing_ml,N_Dosing_ml,P_N_Dosing_Time) %>% filter((!is.na(Phosphorus))|(!is.na(Nitrite))|(!is.na(Ammonia))|(!is.na(P_Dosing_ml))|(!is.na(N_Dosing_ml)))
write.csv(nutrients,"../Data/Nutrient_summary.csv")

# Move "TA" after "pH.total"

pHSlope <- pHSlope %>%
  select(Date, Tank_ID, Time, Date.Time, Observer.Initials, Temperature_C, Salinity_psu, pH.total,Calcium_ppm, TA_dKH, TA, pH_mv, Orion_Temp, light2, light3, light4, tris.date, Probe.Set, notes, `(Intercept)`, TTris, mVTris)
```

## Change to long format

Change data format to long format 
```{r}
pHSlope.long <-pHSlope %>% pivot_longer(cols=Temperature_C:TA,
  names_to = "metric",
  values_to = "value")

nutrients.long <-nutrients %>% pivot_longer(cols=Temperature_C:Ammonia,
  names_to = "metric",
  values_to = "value") %>% filter(!is.na(value)) %>%
  filter(Date >= "2026-04-01")
```

Filter by relevant dates if needed

## Plot

Make a list of dataframes, each containing a horizontal line that will correspond to the upper and lower threshold of each parameter (temperature, salinity, pH total)
```{r}
hlines_data <- list(
  data.frame(yintercept = 25.5, metric = "Temperature_C"), # lower threshold for temperature in C°
  data.frame(yintercept = 26.5, metric = "Temperature_C"), # upper threshold for temperature in C°
  data.frame(yintercept = 34.5, metric = "Salinity_psu"), # lower threshold for salinity in psu
  data.frame(yintercept = 35.5, metric = "Salinity_psu"), # upper threshold for salinity in psu
  data.frame(yintercept = 8.05, metric = "pH.total"), # lower threshold for total pH
  data.frame(yintercept = 8.2, metric = "pH.total"), # upper threshold for total pH
  data.frame(yintercept = 450, metric = "Calcium_ppm"), # upper threshold for Calcium_ppm
  data.frame(yintercept = 300, metric = "Calcium_ppm"), # lower threshold for Calcium_ppm
  data.frame(yintercept = 1800.000, metric = "TA"), # lower threshold for TA
  data.frame(yintercept = 2500.000, metric = "TA") # upper threshold for TA
    )
```

Plot all dates 
```{r}
facet_labels <- unique(pHSlope.long$metric)
names(facet_labels) = facet_labels
facet_labels <- replace(facet_labels, which(facet_labels == "TA"), "TA (µmol kg-1)")
facet_labels <- replace(facet_labels, which(facet_labels == "TA_dKH"), "TA (dKH)")
facet_labels <- replace(facet_labels, which(facet_labels == "Calcium_ppm"), "Calcium (ppm)")
facet_labels <- replace(facet_labels, which(facet_labels == "pH.total"), "pH")
facet_labels <- replace(facet_labels, which(facet_labels == "Salinity_psu"), "Salinity (psu)")
facet_labels <- replace(facet_labels, which(facet_labels == "Temperature_C"), "Temperature (ºC)")

daily_tank<-pHSlope.long %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=2)+
  xlab("Date")+
  facet_grid(factor(metric,c("Calcium_ppm","TA_dKH","TA","pH.total","Salinity_psu","Temperature_C")) ~ ., scales = "free", labeller = as_labeller(facet_labels))+
  geom_hline(data = hlines_data[[1]], aes(yintercept = yintercept), linetype = "dashed") +    
  geom_hline(data = hlines_data[[2]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[3]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[4]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[5]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[6]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[7]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[8]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[9]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[10]], aes(yintercept = yintercept), linetype = "dashed") +
  theme_bw() +
  theme(text = element_text(size = 14)); daily_tank


# Save plot 
ggsave("../Output/Daily_Measurements.pdf", daily_tank, width = 20, height = 15, units = c("in"))
```

It isn't super informative to look at multiple years in one plot. 

Filter and plot by the past six months
```{r}
# Filter to the past 180 days 
daily_6month <- pHSlope.long %>%
  filter(Date >= Sys.Date() - 180)  # Adjust 30 to the number of days you want to consider

# Plot data from the past month
daily_6month_plot <- daily_6month %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(factor(metric,c("Calcium_ppm","TA_dKH","TA","pH.total","Salinity_psu","Temperature_C")) ~ ., scales = "free", labeller = as_labeller(facet_labels))+
  geom_hline(data = hlines_data[[1]], aes(yintercept = yintercept), linetype = "dashed") +    
  geom_hline(data = hlines_data[[2]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[3]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[4]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[5]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[6]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[7]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[8]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[9]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[10]], aes(yintercept = yintercept), linetype = "dashed") +
  theme_bw() +
  theme(text = element_text(size = 14)) 
daily_6month_plot

# Save plot 
ggsave("../Output/Daily_Measurements_Past_6Months.pdf", daily_6month_plot, width = 20, height = 15, units = c("in"))
```

Filter and plot by the past month
```{r}
# Filter to the past 30 days 
daily_month <- pHSlope.long %>%
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

# Plot data from the past month
daily_month_plot <-daily_month %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=4)+
  xlab("Date")+
  facet_grid(factor(metric,c("Calcium_ppm","TA_dKH","TA","pH.total","Salinity_psu","Temperature_C")) ~ ., scales = "free", labeller = as_labeller(facet_labels))+
  geom_hline(data = hlines_data[[1]], aes(yintercept = yintercept), linetype = "dashed") +    
  geom_hline(data = hlines_data[[2]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[3]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[4]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[5]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[6]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[7]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[8]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[9]], aes(yintercept = yintercept), linetype = "dashed") +
  geom_hline(data = hlines_data[[10]], aes(yintercept = yintercept), linetype = "dashed") +
  theme_bw() +
  theme(text = element_text(size = 14)) + scale_x_date(date_breaks = "1 day", date_labels = "%a\n%b %d")
daily_month_plot

# Save plot 
ggsave("../Output/Daily_Measurements_Past_Month.pdf", daily_month_plot, width = 20, height = 15, units = c("in"))
```

Summarize daily measurements over the past month
```{r}
daily_month <- pHSlope %>% 
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

summary<-daily_month%>%
  group_by(Tank_ID)%>%
  #select(!tank)%>%
  select(Temperature_C:TA) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary
```


## Plot Nutrients

```{r}
nutrient_dosing <- nutrients.long %>% filter(!is.na(P_Dosing_ml)| !is.na(N_Dosing_ml)) %>%
  distinct(Date, Tank_ID, P_Dosing_ml, N_Dosing_ml, P_N_Dosing_Time) %>% mutate(
    dosing_info = paste0(Tank_ID, "\ndosed\n","", P_Dosing_ml, " ml P & ", N_Dosing_ml, " ml N\n","@", P_N_Dosing_Time))
```


Plot all dates 
```{r}
facet_labels <- unique(nutrients.long$metric)
names(facet_labels) = facet_labels
facet_labels <- replace(facet_labels, which(facet_labels == "Phosphorus"), "Phosphorus (ppb)")
facet_labels <- replace(facet_labels, which(facet_labels == "Nitrite"), "Nitrite (ppb NO2-)")
facet_labels <- replace(facet_labels, which(facet_labels == "Ammonia"), "Ammonia (ppm NH3)")
facet_labels <- replace(facet_labels, which(facet_labels == "Temperature_C"), "Temperature (ºC)")

daily_tank<-nutrients.long %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=2)+
  geom_vline(data = nutrient_dosing,aes(xintercept = as.numeric(Date)),linetype = "dashed",linewidth = 0.7,alpha = 0.6) +
  geom_label_repel(data = nutrient_dosing %>% crossing(metric = c("Phosphorus")),
    aes(x = Date,y = Inf,label = dosing_info),size = 3,
    segment.colour = "grey30",segment.size = 0.4,nudge_x = 0.3, nudge_y = -10,
    label.padding = unit(0.15, "lines"),show.legend = FALSE,inherit.aes = FALSE) +
  xlab("Date")+
  facet_grid(factor(metric,c("Phosphorus","Nitrite","Ammonia","Temperature_C")) ~ ., scales = "free", labeller = as_labeller(facet_labels))+
  theme_bw() +
  theme(text = element_text(size = 14)); daily_tank

# Save plot 
ggsave("../Output/Daily_Nutrients.pdf", daily_tank, width = 10, height = 8, units = c("in"))
```

Filter and plot by the past month
```{r}
# Filter to the past 30 days 
daily_month <- nutrients.long %>%
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

nutrient_dosing_month <- nutrient_dosing %>%
  filter(Date >= Sys.Date() - 30) 

# Plot data from the past month
daily_month_plot <-daily_month %>%
  ggplot(aes(x=Date, y=value, colour=Tank_ID))+
  geom_point(size=2)+
  geom_vline(data = nutrient_dosing_month,aes(xintercept = as.numeric(Date)),linetype = "dashed",linewidth = 0.7,alpha = 0.6) +
  geom_label_repel(data = nutrient_dosing_month %>% crossing(metric = c("Phosphorus")),
    aes(x = Date,y = Inf,label = dosing_info),size = 3,
    segment.colour = "grey30",segment.size = 0.4,nudge_x = 0.3, nudge_y = -10,
    label.padding = unit(0.15, "lines"),show.legend = FALSE,inherit.aes = FALSE) +
  xlab("Date")+
  facet_grid(factor(metric,c("Phosphorus","Nitrite","Ammonia","Temperature_C")) ~ ., scales = "free", labeller = as_labeller(facet_labels))+
  theme_bw() +
  theme(text = element_text(size = 14)) + scale_x_date(date_breaks = "1 day", date_labels = "%a\n%b %d")
daily_month_plot

# Save plot 
ggsave("../Output/Daily_Nutrients_Past_Month.pdf", daily_month_plot, width = 10, height = 8, units = c("in"))
```

Summarize daily measurements over the past month
```{r}
daily_month <- nutrients %>% 
  filter(Date >= Sys.Date() - 30)  # Adjust 30 to the number of days you want to consider

summary<-daily_month%>%
  group_by(Tank_ID)%>%
  #select(!tank)%>%
  select(Temperature_C:Ammonia) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary
```

## TA-dkh

```{r}
library(ggrepel)
library(ggpmisc)

pHSlope %>% group_by(Tank_ID,Date,TA_dKH) %>% drop_na(TA) %>% summarize(TA = mean(TA)) %>%
  ggplot(aes(x = TA, y = TA_dKH)) +
  geom_point(size = 2) + geom_text_repel(aes(label=TA_dKH)) + 
  geom_smooth(method = "lm", se = FALSE) +  # use linear model (lm) and remove shaded error band
  theme_minimal() 

pHSlope$Date <- as.factor(pHSlope$Date)

TA_plot <- pHSlope %>% 
  group_by(Tank_ID, Date, TA_dKH) %>% 
  drop_na(TA) %>% 
  summarize(TA = mean(TA), .groups = "drop") %>% drop_na %>% droplevels() %>%
  ggplot(aes(x = TA, y = TA_dKH, color=Date)) +
  geom_point(size = 2) + geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_text_repel(aes(label = paste0("TA=", round(TA, 1), "\nTA_dKH=", round(TA_dKH, 1)))) +
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y ~ x,
               parse = TRUE,
               label.x = "left",
               label.y = "top",color="black") +
  theme_minimal()+
  labs(
    title = "Relationship Between TA and TA_dKH in Blue Tanks",
    x = "Total Alkalinity (TA)",
    y = "TA in dKH"
  )

ggsave("../Output/TA_dKH_relationship.pdf", TA_plot, width = 7, height = 5, units = c("in"))
```


#PAM Data for Aquarium Room- Adapted from ZD's PAM Script for Timeseries (https://github.com/zdellaert/TimeSeries/blob/main/3-Pacu/code/PAM.md)
```{r}
### 1. PAM Data Fv/fm
PAM <- read.csv("../Data/PAM_CBLS_Aquarium.csv") %>% clean_names
PAM <- PAM %>% 
  filter(tank_id == "BT1/2" | tank_id == "BT3/4")%>%
  mutate(date = as.factor(date)) %>% 
                mutate(timepoint = factor(timepoint)) %>%#,levels = c("0","1","3","6","12",
                                                    #           "24","36","48","72","120","170"), ordered = TRUE)) %>% 
                mutate(treatment = factor(coral_species,levels = c("PCOM","MCAP","PACU","ELAM", ordered=TRUE)))

PAM_fv_fm <- ggplot(PAM,aes(x = date, y = fv_fm_y_1000)) + 
    geom_boxplot(aes(fill = coral_species)) + labs(x = "Date", y = "Fv/Fm", title = "Fv/Fm by Date and Tank") +
  theme_minimal()

ggsave("../Output/PAM_Fv_Fm.pdf", PAM_fv_fm, width = 7, height = 5, units = c("in"))


#state compare means
PAM_f <- ggplot(PAM,aes(x = coral_species, y = fv_fm_y_1000)) + 
    geom_boxplot(aes(fill = coral_species)) + labs(x = "Date", y = "Fv/Fm", title = "Fv/Fm by Date and Treatment") +
    facet_grid(~timepoint )+
  theme_minimal()+
  stat_compare_means(
  label = "p.signif"
)

PAM_f

ggsave("../Output/PAM_f.pdf", plot = last_plot(), width = 9, height = 4, bg = "white")


#anovamodel  
PAM_anova <- ggplot(PAM,aes(x = coral_species, y = fv_fm_y_1000)) + 
    geom_boxplot(aes(fill = date)) + labs(x = "Coral Species", y = "Fv/Fm", title = "Fv/Fm by Date and Coral Species") +
  theme_minimal()+
  stat_compare_means(aes(group = date), method = "anova", label = "p.format", size=3)

PAM_anova
ggsave("../Output/PAM_anova.pdf", plot = last_plot(), width = 9, height = 4, bg = "white")
```

#Water quality checker
##Plot 1
#```{r}
# Read in data and save output
pdf("../Output/Nutrients_Quality_Testing_Plot.pdf", width = 7, height = 5)


nutrient_checker <- read.csv("../Data/Nutrient_Quality_Checker.csv")
head(nutrient_checker)

#box checker plot for water testing checker
boxplot(nutrient_checker[, c("Ammonia_ppm", "dKH", "Nitrite_ppb", "Nitrate_ppm", "Phosphorus_ppb")],
        main = "Water Nutrients Quality Testing Plot",
        xlab = "Nutrient Type",
        ylab = "Concentration in ppm/ppb",
        col = c("skyblue", "lightgreen", "salmon", "plum", "gold"),
        names = c("Ammonia", "dKH", "Nitrite", "Nitrate", "Phosphorus"),
        las = 1)

dev.off()
```








#```{r}
library(ggplot2)
library(tidyr)

data_long <- pivot_longer(nutrient_checker,
                          cols = c("Ammonia_ppm", "dKH", "Nitrite_ppb", "Nitrate_ppm", "Phosphorus_ppb"),
                          names_to = "Assay",
                          values_to = "Value")

p <- ggplot(data_long, aes(x = "", y = Value, fill = Assay)) +
  geom_boxplot(width = 0.4, show.legend = FALSE) +
  facet_wrap(~Assay, scales = "free_y", ncol = 3) +

  # Clean theme
  theme_classic() +

  labs(title = "Water Measurement by Nutrients",
       y = "Concentration",
       x = NULL) +

  theme(
    strip.text = element_text(size = 12, face = "bold"),   # facet titles
    axis.text.x = element_blank(),                         # remove useless x labels
    axis.ticks.x = element_blank(),
    panel.spacing = unit(1.2, "lines"),                    # space between plots
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

p

ggsave("../Output/Nutrient_checker_Faceted.pdf", p, width = 7, height = 5, units = c("in"))
```

##Plot 2
#```{r}
library(ggplot2)
library(tidyr)

# Convert to long format
data_long <- pivot_longer(nutrient_checker,
                          cols = c("Ammonia_ppm", "dKH", "Nitrite_ppb", "Nitrate_ppm", "Phosphorus_ppb"),
                          names_to = "Assay",
                          values_to = "Value")

# Faceted boxplot
p <- ggplot(data_long, aes(x = Assay, y = Value, fill = Assay)) +
  geom_boxplot() +
  facet_wrap(~Assay, scales = "free_y") +
  theme_minimal() +
  labs(title = "Water tests by Nutrients",
       x = "",
       y = "Concentration") +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

ggsave("../Output/Nutrient_checker_Faceted.pdf", p, width = 7, height = 5, units = c("in"))
```

